---
title: "Homework 4"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http://cran.rstudio.com"))
```



## HPC
### Problem 1: Make sure your code is nice

#### Create a matrix with a set seed
```{r}
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)
```
#### Find total row sums
```{r}
fun1 <- function(dat) {
  n <- nrow(dat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(dat[i, ])
  }
  ans
}
```
#### Alternative code to run faster
```{r}
fun1alt <-function(dat) {
  x <- rowSums(dat)
  return(x)
}
```

#### Find cumulative row sums

```{r}
# Cumulative sum by row
fun2 <- function(dat) {
  n <- nrow(dat)
  k <- ncol(dat)
  ans <- dat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- dat[i, j] + ans[i, j - 1]
    }
  }
  ans
}
```

#### Alternative code to run faster
```{r}

fun2alt <- function(mat) {
    t(apply(mat, 1, cumsum))}

```

### Benchmarking
#### Test for the first
```{r}
summary(microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), check = "equivalent"
), unit="relative")

```

#### Test for the second
```{r}
summary(microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
), unit="relative")
```

### Problem 2. Make things run faster with parallel computing

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}


# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132

# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```



#### Use parlapply to run faster 
```{r}
library(parallel)


cl <- makePSOCKcluster(4L)
clusterSetRNGStream(cl, 1231)
system.time({
  clusterExport(cl, "sim_pi")
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n=10000))
  print(mean(ans))

})
stopCluster(cl)

```




## SQL

### Setup
```{r}
if (!require(RSQLite)) install.packages(c("RSQLite"))
if (!require(DBI))     install.packages(c("DBI"))
library(RSQLite)
library(DBI)



# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)

```

```{r}
dbListTables(con)
```

### Brief data exploration to determine variable names

```{r}
#View variables in each dataset
dbGetQuery(con, "PRAGMA table_info(film)")
dbGetQuery(con, "PRAGMA table_info(film_category)")
dbGetQuery(con, "PRAGMA table_info(category)")
```

### Question 1. How many movies are avaliable in each rating category?
```{r}
dbGetQuery(con, "
SELECT rating, COUNT (*) AS Rating_Count
FROM film
GROUP by rating
ORDER BY rating")
```

### Question 2. What is the average replacement cost and rental rate for each rating category?
```{r}
dbGetQuery(con, "
SELECT rating, 
       AVG(replacement_cost) AS Avg_Replacement_Cost,
       AVG(rental_rate) AS Avg_Rental_Rate
FROM film
GROUP by rating
ORDER BY rating")
```

### Question 3. Use table film_category together with film to find how many films there are with each category ID.
```{r}
dbGetQuery(con, "
SELECT category_id, COUNT (category_id) AS Film_Count 
FROM (SELECT c.film_id, c.category_id, p.film_id 
FROM film_category AS c INNER JOIN film AS p
ON c.film_id = p.film_id)
GROUP BY category_id
ORDER BY Film_Count DESC")
```

### Question 4. Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{r}
dbGetQuery(con, "
SELECT category_id, name, COUNT (category_id) AS Film_Count 
FROM (SELECT a.name, a.category_id, b.category_id, b.film_id
FROM (SELECT c.film_id, c.category_id, p.film_id 
FROM film_category AS c INNER JOIN film AS p
ON c.film_id = p.film_id) AS b INNER JOIN category AS a 
ON a.category_id = b.category_id)
GROUP BY category_id
ORDER BY Film_Count DESC")
```

#### _Sports seems to be the most popular film category!_

#### Disconnect
```{r}
dbDisconnect(con)
```






