---
title: "Lab 10"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "http//cran.rstudio.com"))
```

## R Markdown
```{r}
library(RSQLite)
library(DBI)
```
```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
```
```{r}
dbListTables(con)
```

```{sql, connection=con}
PRAGMA table_info(actor)
```

```{r}
#in order to do the SQL query FROM R use the dbGetQuery command to do it
dbGetQuery(con, "PRAGMA table_info(actor)")
```
## Exercise 1

```{r}
dbGetQuery(con, "
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
LIMIT 20")
```
```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by last_name, first_name
LIMIT 15
```

## Exercise 2.
```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
ORDER BY last_name
```
## Exercise 3. 

```{r}
dbGetQuery(con,"
      PRAGMA table_info(rental)
           "
)
```


```{r}
dbGetQuery(con,"
SELECT DISTINCT customer_id, rental_date
FROM  rental
WHERE date(rental_date) = '2005-07-05'
")
```


```{r}
dbGetQuery(con,"
SELECT customer_id, 
     COUNT(*) AS N
FROM  rental
WHERE date(rental_date) = '2005-07-05'
GROUP BY customer_id
")
```

## Exercise 4.1 
Construct a query that retrieves all rows from the payment table where the amount is either 1.99, 7.99, 9.99.

```{r}
dbGetQuery(con,"
      PRAGMA table_info(payment)
")
```

```{r}
dbGetQuery(con,"
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
LIMIT 10
  ")
```

## Exercise 4.2

Construct a query that retrieves all rows from the payment table where the amount is greater than 5

```{r}
dbGetQuery(con,"
SELECT *
FROM payment
WHERE amount > 5
LIMIT 10
")
```

## Exercise 4.2, second part --Greater than 5 and less than 8

```{r}
dbGetQuery(con,"
SELECT *
FROM payment
WHERE amount > 5   AND amount < 8
LIMIT 10
")
```
## Exercise 5. Retrieve all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.

```{r}
dbGetQuery(con,"
SELECT c.customer_id, c.last_name, p.payment_id, p.amount
FROM customer AS c INNER JOIN payment AS p
ON c.customer_id = p.customer_id
WHERE c.last_name=='DAVIS'
/* WHERE c.last_name == 'DAVIS' */ /* This is a comment */
")
```

## Exercise 6. Use COUNT(*) to count the number of rows in rental

```{r}
dbGetQuery(con,"
SELECT COUNT(*)
FROM rental")
```
## Exercise 6.2 - Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id
```{r}
dbGetQuery(con,"
SELECT customer_ID, COUNT(*) as count
FROM rental
GROUP BY customer_id
LIMIT 10")

```

## Exercise 6.3 - Repeat the previous query and sort by the count in descending order
```{r}
dbGetQuery(con,"
SELECT customer_ID, COUNT(*) as count
FROM rental
GROUP BY customer_id
ORDER BY count DESC
LIMIT 10")
```

## Exercise 6.4 - Repeat the previous query but use HAVING to only keep the groups with 40 or more.
```{r}
dbGetQuery(con,"
SELECT customer_ID, COUNT(*) as count
FROM rental
GROUP BY customer_id
HAVING count >= 40
ORDER BY count DESC
LIMIT 10
")
```
## Exercise 7.1 - Modify the above query to do those calculations for each customer_id

```{r}
dbGetQuery(con,"
SELECT customer_ID, 
      COUNT(*) as N,
       MAX(amount) AS maxpayment,
      MIN(amount) AS minpayment,
      AVG(amount) AS avgpayment,
      SUM(amount) AS sumpayment
      
FROM payment
GROUP BY customer_id
ORDER BY N DESC
LIMIT 10
")
```
## Exercise 7.2 - Modify the above query to only keep the customer_ids that have more then 5 payments
```{r}
dbGetQuery(con,"
SELECT customer_ID, 
      COUNT(*) as N,
       MAX(amount) AS maxpayment,
      MIN(amount) AS minpayment,
      AVG(amount) AS avgpayment,
      SUM(amount) AS sumpayment
      
FROM payment
GROUP BY customer_id
HAVING N >5
ORDER BY N DESC
")
```

```{r}
# clean up
dbDisconnect(con)
```

