---
title: "Homework 2"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(tidyverse)
library(dtplyr)
library(knitr)
library(leaflet)
library(ggplot2)
```
#Load in data from github
```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "individual.gz", method="libcurl", timeout = 60)
individual <- data.table::fread("individual.gz")
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "regional.gz", method="libcurl", timeout = 60)
regional <- data.table::fread("regional.gz")
  
```
```{r}
dim(individual)
head(individual)
```
```{r}
head(regional)
#Trim down this dataset so it just has the info we are going to use (PM2.5 data and location for the leaflet)
regional_trim <- regional[, c("townname", "pm25_mass", "lon", "lat")]
regional_trim
```
#How many unique townnames are there? 
```{r}
unique(individual$townname)
unique(regional_trim$townname)
```
#There are 12 unique towns

#Check variables for merging. The 12 unique townnames match with the 12 rows from the regional dataset. 

#Merge data
```{r}
hw2data <-merge(individual, regional_trim, by="townname" )
dim(hw2data)
```

#_This looks correct, there are 1200 rows and now the three additional columns added in the regional (trimmed) dataset
```{r}
head(hw2data)
tail(hw2data)
```
#Which variables have missing values?
```{r}
colSums(is.na(hw2data))
```
#_Varibles that have missing values that we will need for our analyses are: BMI (89/1200 =7.4% missing), FEV (95/1200 = 7.9% missing), Smoke (40/1200 = 3.3% missing), Gas Stove (33/1200 = 2.8% missing), Asthma (31/1200=2.58% missing)

#Calculate all the imputation percentages with code and put them in a table

```{r}
Table_imputation_Percents = hw2data %>% 
    summarise(ASTHMA_percent_imputed = sum(is.na(hw2data$asthma))/n()*100,
              BMI_percent_imputed = sum(is.na(hw2data$bmi))/n()*100,
              FEV_percent_imputed = sum(is.na(hw2data$fev))/n()*100,
              SMOKE_percent_imputed = sum(is.na(hw2data$smoke))/n()*100,
              GASSTOVE_percent_imputed = sum(is.na(hw2data$gasstove))/n()*100)

Table_imputation_Percents <-as.data.frame(Table_imputation_Percents)
colnames(Table_imputation_Percents) <- c("Asthma", "BMI", "FEV", "Smoke", "Gas Stove")
                                        
knitr::kable(Table_imputation_Percents, align=c("c", "c", "c", "c","c"), digits =2, caption = "Imputated Percentages by Variable")

```
#_We can see that all the imputation percentages are pretty low, below 10%. Asthma, smoke, and gasstove are even below 5%._ 

## Data Wrangling Step 1


#For missing values, impute data using the average within the variables “male” and “hispanic.”
```{r}
#Impute variables with missing values grouped by sex/hispanic
hw2data[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = T)),
    by = .(male, hispanic)]
hw2data[, fev_imp := (fcoalesce(fev, mean(fev, na.rm = T))),
    by = .(male, hispanic)]

#Create another smoke, asthma, and gasstove variable that is numeric so it can be imputed
hw2data [ , smoke_num := (as.numeric(smoke, na.rm=F))]
hw2data [ , gasstove_num := (as.numeric(gasstove, na.rm=F))]
hw2data [ , asthma_num := (as.numeric(asthma, na.rm=F))]

#Now try imputation again
hw2data[, smoke_imp := (fcoalesce(smoke_num, median(smoke_num, na.rm = T))),
    by = .(male, hispanic)]
hw2data[, gasstove_imp := (fcoalesce(gasstove_num, median(gasstove_num, na.rm = T))),
    by = .(male, hispanic)]
hw2data[, asthma_imp := (fcoalesce(asthma_num, median(asthma_num, na.rm = T))),
    by = .(male, hispanic)]
```

#Make sure that the imputed variables do not have any NA values - specifically we are checking bmi_imp, fev_imp, smoke_imp, gasstove_imp, asthma_imp 
```{r}
colSums(is.na(hw2data))
```

#_This worked. None of the imputed variables have missing data._

## Step 2 

#Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r}
hw2data[, obesity_level := fifelse(bmi_imp < 14, "Underweight", 
                fifelse(bmi_imp >= 14 & bmi_imp <22, "Normal",
                fifelse(bmi_imp >=22 & bmi_imp <24, "Overweight", "Obese")))]
#Make sure categories look ok
table(hw2data$obesity_level)
```
#These look appropriate

#Make table with max, min , and total number observations per category.
```{r}
hw2data %>% 
  group_by(obesity_level) %>% 
          summarise(n(),
          Min_BMI = min(bmi_imp),
          Max_BMI = max(bmi_imp)) %>% 
              arrange(Max_BMI)
```
## Step 3 

#Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r}
hw2data[, smoke_gas_exposure := fifelse(smoke_imp ==0 & gasstove_imp ==0, "None", 
                fifelse(smoke_imp ==1 & gasstove_imp ==0, "Smoke Only",
                fifelse(smoke_imp ==0 & gasstove_imp ==1, "Stove Only", "Both")))]
table(hw2data$smoke_gas_exposure)
```
## Step 4 - Create Summary tables

#Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r}
#Table 1
Table_Town = hw2data %>% 
  group_by(townname) %>% 
          summarise(n = n(), sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))
Table_Town <- as.data.frame(Table_Town)
colnames(Table_Town) <- c("Town Name","N", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")
knitr::kable(Table_Town, align=c("l", "c", "c", "c","c"), digits =2, caption = "Table 1. Forced expiratory volume in 1 second (mL) (FEV1) by Town")
```
#_Atascadero has much higher rates of asthma compared to other cities. The next highest would be San Dimas and then Lancaster. It looks like the FEV1 for these cities is right around the average for the dataset as a whole (they aren't especially low despite their high asthma percentages)._
```{r}
#Table 2
#Make a labeled version of male and female so its easy to work with
sex <- as.factor(hw2data$male)
hw2data[, sex := fifelse(male ==0 , "Female", "Male")]
Table_Sex = hw2data %>% 
  group_by(sex) %>% 
          summarise(n = n(), sum(asthma_imp==1)/n()*100,
                    mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))
       Table_Sex <- as.data.frame(Table_Sex)
colnames(Table_Sex) <- c("Sex", "N", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")
knitr::kable(Table_Sex, align=c("l", "c", "c", "c","c"), digits =2, caption = "Table 2. Forced expiratory volume in 1 second (mL) (FEV1) by Sex")
```
#_Males have a higher risk of asthma comapred to females in this dataset, 17% versus 12%. Males have a higher FEV1 and this is likely due to them just being larger and having bigger lungs capable of larger expiratory volumes._

```{r}
#Table 3
Table_Obesity = hw2data %>% 
  group_by(obesity_level) %>% 
          summarise(n = n(), sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))
Table_Obesity <- as.data.frame(Table_Obesity)
colnames(Table_Obesity) <- c("Obesity Level", "N", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")
knitr::kable(Table_Obesity, align=c("l", "c","c", "c","c"), digits =2, caption = "Table 3. Forced expiratory volume in 1 second (mL) (FEV1) by Obesity Level")
```

#_Obesity level does appear to be associated with presence of asthma. As BMI gets higher, prevalence of asthma increases. The rate of asthma between normal weight and obese subjects is hugely different, an almost 50% increase in risk. The FEV1 appears to go up as BMI goes up. This is counter intuitive to what we see with asthma risk (since asthma reduces FEV1), but I suspect that FEV1 going up as BMI goes up is a function of larger children with larger lungs having higher FEV1s due to body size and overall development._

```{r}
#Table 4
Table_Smoke_Gas_Exp = hw2data %>% 
  group_by(smoke_gas_exposure) %>% 
          summarise(n = n(), sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))
Table_Smoke_Gas_Exp <- as.data.frame(Table_Smoke_Gas_Exp)
colnames(Table_Smoke_Gas_Exp) <- c("Smoke Gas Exposure", "N","Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")
knitr::kable(Table_Smoke_Gas_Exp, align=c("l", "c", "c", "c","c"), digits =2, caption = "Table 4. Forced expiratory volume in 1 second (mL) (FEV1) by Second Hand Smoke and Gas Stove Exposure")
```
#_The highest rate of asthma is in patients with only second hand smoke exposure. Though the N for this group is very low (only 36). The lowest asthma percent is in the both group which is a bit confusing. Mean FEV1 is highest in the none and smoke only groups, with the Stove only and both groups having the lowest FEV1. This suggests that the gas stove is the exposure that is most likely associated with a lower FEV1._

## Data Visualization

##Step 1 FEV by BMI grouped by townname
```{r}
hw2data%>% 
  ggplot(mapping = aes(x = bmi_imp, y = fev_imp)) + 
  geom_point(mapping = aes(color = townname)) + 
  scale_color_brewer(palette="Paired")+
  geom_smooth(method = lm, mapping = aes(linetype = townname)) +
  facet_wrap(~ townname, nrow = 3) +
  labs(title="Forced Expiratory Volume by Body Mass Index", x="Body Mass Index (kg/m2)", y= "Forced Expiratory Volume in 1 Second (mL)")
```

#_Overall the BMI and FEV1 variables appear to have a positive association. As BMI increases, FEV1 increases. This is true across all 12 towns. Some towns have a more dramatic slope than others, especially Atascadero, Lake Elsinore, and Upland. In these cities it looks like the data have fewer high BMI subjects pulling the FEV line down (like we see in Riverside for instance)._

##Step 2 - Make stacked histograms 
```{r}

hw2data %>%
  ggplot(mapping =aes(x=fev_imp, fill = obesity_level, color = obesity_level)) +   
  geom_histogram(bin = 100, binwidth = 100, color="black")+
  labs(title="Forced Expiratory Volume by Body Mass Index", x="Forced Expiratory Volume in 1 second (mL)", y= "Observations")+
scale_fill_brewer(palette = "Set2")
```

#_The obese and overweight bars (orange and blue) appear to have their own mean that is pushed toward a higher FEV1 volume compared to the overall mean of the data. Conversely the underweight subjects (pink) appear to have a mean on the lower end of the FEV1 spectrum. This is a nice visual of what we saw in the table from a previous question._

```{r}

hw2data %>%
  ggplot(mapping =aes(x=fev_imp, fill = smoke_gas_exposure, color = smoke_gas_exposure)) +   
  geom_histogram(bin = 100, binwidth = 100, color="black" )+
  labs(title="Forced Expiratory Volume by Second Hand Smoke & Gas Stove Exposure", x="Forced Expiratory Volume in 1 second (mL)", y= "Number of Observations")+
scale_fill_brewer(palette = "Set1")
```

#_The majority of patients in this dataset have exposure to a gas stove (purple, red). The sheer numer of stove exposures make it harder to tease out whether smoke or gas stove are independently associated with FEV, at least in this graphical view. I think stratifying the data between gas stove or no gas stove and then evaluating again would be more fruitful. However there does appear to be more exposure to both in the lower end of FEV around 1600mL where the red section appears to be a higheer proportion of the total bar versus the rest of the graph._

## Step 3 - Make bar chart of BMI by smoke/gas exposure
```{r}
hw2data%>%
  ggplot(mapping = aes(x = smoke_gas_exposure, y = bmi_imp, fill = smoke_gas_exposure)) + 
    stat_summary(fun.y =  mean, geom = "bar") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar", width = 0.2) +
  stat_summary(fun.data = mean_sdl, geom = "point")+
labs(title="Body Mass Index by Second Hand Smoke and Gas Stove Exposure", x="Second Hand Smoke and Gas Stove Exposure", y= "Mean Body Mass Index (kg/m2)")+
scale_fill_brewer(palette = "Blues")+
  theme_get()

```

#_It looks as though BMI is higher in patients with Smoke only exposure or both Smoke and Gas Stove exposure. This suggests that second hand smoke exposure is likely the more causative (or more highly associated) factor when it comes to these exposures and BMI. The effects of a gas stove appear minimal as the mean BMI for Stove Only looks very similar to no exposure at all._

```{r}
hw2data %>% 
  ggplot(mapping = aes(x = bmi_imp, y = fev_imp)) + 
  geom_point(color="blue") + 
  geom_smooth(method = lm, color="red")+
labs(title="Forced Expiratory Volume by Body Mass Index", x="Body Mass Index(kg/m2)", y= "Forced Expiratory Volume in 1 second (mL)")
```

#_The positive association between FEV1 and BMI is probably due to larger children having bigger lungs and higher FEV1 values for their age. We know from the table from a previous question that higher BMI is associated with higher rates of asthma. So if we added asthma as a variable to this evaluation it would probably be useful for seeing how BMI interacts with FEV and asthma._ 

```{r}
hw2data %>% 
  ggplot(mapping = aes(x = smoke_gas_exposure, y = fev_imp, fill=smoke_gas_exposure)) + 
  geom_boxplot(color="black") +
labs(title="Forced Expiratory Volume by Second Hand Smoke and Gas Stove Exposure", x="Second Hand Smoke and Gas Stove Exposure", y= "Forced Expiratory Volume in 1 second (mL)")+
  scale_fill_brewer(palette = "Pastel1")
```

#_Smoke and gas exposure may have a small effect on FEV1. The effect is probably coming from the stove exposure. The means for Both and Stove Only are very close, and the means for None and Smoke Only are very close. We know from the table that the difference in FEV1 is about 30 (out of around 2000mL), or a change of 1-2%. This is pretty small. One would probably need a larger dataset with other variables controlled for in order to demonstrate the relationship between these variables._

## Step 5 - Create leaflet plot
```{r}
# Generating a color palette
pal1 <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=hw2data$pm25_mass)
pal1
```
```{r}
prettymap <- leaflet(hw2data) %>% 
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~lat, lng=~lon,
    label = ~paste0(pm25_mass), color = ~ pal1(pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  # And legend
  addLegend('bottomleft', pal=pal1, values=hw2data$pm25_mass,
          title='PM 2.5 Concentration by City', opacity=1)

prettymap
```

#_The cities examined are in central and southern California. It looks like there is worse air quality (higher PM2.5 concentrations) near the more urban areas like Los Angeles and Long Beach (the port is a source of pollution). This makes sense, more cars, more planes, more sources of air pollution._

## Step 6 - Choose a visualization to examine whether PM2.5 mass is associated with FEV.

#Make graph using a regression line to see what kind of relationship there is between PM2.5 and FEV1

```{r}
hw2data %>% 
  ggplot(mapping = aes(x = pm25_mass, y = fev_imp)) + 
  geom_point(color="blue") + 
  geom_smooth(method = lm, color="purple")+
labs(title="Forced Expiratory Volume by PM2.5 Concentration", x="PM2.5 Concentration", y= "Forced Expiratory Volume in 1 second (mL)")
```

#_There appears to be a negative relationship between FEV1 and PM2.5 concentration. The regression line (purple) slopes downward. In other words as PM2.5 concentration increases, FEV1 decreases. This makes sense as PM2.5 is known to be an air pollutant that harms the lungs. This data would suggest that PM2.5 is associated with reduced lung function (lower FEV1 indicates worse lung function) though we are not looking at statistical significance so this trend may not be statistically significant._

