---
title: "Homework 2"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(tidyverse)
library(dtplyr)
library(knitr)
library(leaflet)
library(ggplot2)
```
#Load in data from github
```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "individual.gz", method="libcurl", timeout = 60)

individual <- data.table::fread("individual.gz")

download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "regional.gz", method="libcurl", timeout = 60)

regional <- data.table::fread("regional.gz")
  
```
```{r}
dim(individual)
head(individual)
```
```{r}
head(regional)
#Trim down this dataset so it just has the info we are going to use (PM2.5 data and location for the leaflet)
regional_trim <- regional[, c("townname", "pm25_mass", "lon", "lat")]
regional_trim

```
#How many unique townnames are there? 
```{r}
unique(individual$townname)
unique(regional_trim$townname)
```

#Check variables for merging. There are 12 unique townnames. This matches with the 12 rows from the regional dataset. 
#Merge data
```{r}
hw2data <-merge(individual, regional_trim, by="townname" )
dim(hw2data)
```

#_This looks correct, there are 1200 rows and now the three additional columns added in the regional (trimmed) dataset
```{r}
head(hw2data)
tail(hw2data)
```
#Which variables have missing values?
```{r}
colSums(is.na(hw2data))
```

## Data Wrangling Step 1



#For missing values, impute data using the average within the variables “male” and “hispanic.” For variables with 0/1 values will impute with the median.

```{r}
#Impute variables with missing values grouped by sex/hispanic

hw2data[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = T)),
    by = .(male, hispanic)]


hw2data[, fev_imp := (fcoalesce(fev, mean(fev, na.rm = T))),
    by = .(male, hispanic)]

#Create another smoke, asthma, and gasstove variable that is numeric so it can be imputed

hw2data [ , smoke_num := (as.numeric(smoke, na.rm=F))]
hw2data [ , gasstove_num := (as.numeric(gasstove, na.rm=F))]
hw2data [ , asthma_num := (as.numeric(asthma, na.rm=F))]

#Now try imputation again

hw2data[, smoke_imp := (fcoalesce(smoke_num, median(smoke_num, na.rm = T))),
    by = .(male, hispanic)]

hw2data[, gasstove_imp := (fcoalesce(gasstove_num, median(gasstove_num, na.rm = T))),
    by = .(male, hispanic)]

hw2data[, asthma_imp := (fcoalesce(asthma_num, median(asthma_num, na.rm = T))),
    by = .(male, hispanic)]

```
```{r}
head(hw2data)
```

# This worked. Now my dataset has the imputed variables in place. 

## Step 2 

#Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.

```{r}
hw2data[, obesity_level := fifelse(bmi_imp < 14, "Underweight", 
                fifelse(bmi_imp >= 14 & bmi_imp <22, "Normal",
                fifelse(bmi_imp >=22 & bmi_imp <24, "Overweight", "Obese")))]

#Make sure categories look ok
table(hw2data$obesity_level)

```
# These look appropriate
# Make table with max, min , and total number observations per category
```{r}
hw2data %>% 
  group_by(obesity_level) %>% 
          summarise(n(),
          Min_BMI = min(bmi_imp),
          Max_BMI = max(bmi_imp)) %>% 
              arrange(Max_BMI)
```
## Step 3 
# Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.

```{r}
hw2data[, smoke_gas_exposure := fifelse(smoke_imp ==0 & gasstove_imp ==0, "None", 
                fifelse(smoke_imp ==1 & gasstove_imp ==0, "Smoke Only",
                fifelse(smoke_imp ==0 & gasstove_imp ==1, "Stove Only", "Both")))]

table(hw2data$smoke_gas_exposure)
```
## Step 4 Create Summary tables

# Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”

```{r}
#Table 1
Table_Town = hw2data %>% 
  group_by(townname) %>% 
          summarise(sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))


Table_Town <- as.data.frame(Table_Town)


colnames(Table_Town) <- c("Town Name", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")


knitr::kable(Table_Town, align=c("l", "c", "c","c"), digits =2, caption = "Table 1. Forced expiratory volume in 1 second (mL) (FEV1) by Town")

```

```{r}
#Table 2

#Make a labeled version of male and female so its easy to work with

sex <- as.factor(hw2data$male)

hw2data[, sex := fifelse(male ==0 , "Female", "Male")]


Table_Sex = hw2data %>% 
  group_by(sex) %>% 
          summarise(sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))

Table_Sex <- as.data.frame(Table_Sex)


colnames(Table_Sex) <- c("Sex", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")


knitr::kable(Table_Sex, align=c("l", "c", "c","c"), digits =2, caption = "Table 2. Forced expiratory volume in 1 second (mL) (FEV1) by Sex")

```

```{r}
#Table 3

Table_Obesity = hw2data %>% 
  group_by(obesity_level) %>% 
          summarise(sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))


Table_Obesity <- as.data.frame(Table_Obesity)


colnames(Table_Obesity) <- c("Obesity Level", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")

knitr::kable(Table_Obesity, align=c("l", "c", "c","c"), digits =2, caption = "Table 3. Forced expiratory volume in 1 second (mL) (FEV1) by Obesity Level")
```
```{r}
#Table 4

Table_Smoke_Gas_Exp = hw2data %>% 
  group_by(smoke_gas_exposure) %>% 
          summarise(sum(asthma_imp==1)/n()*100,
          mean_fev  = mean(fev_imp),
          SD_fev = sd(fev_imp))


Table_Smoke_Gas_Exp <- as.data.frame(Table_Smoke_Gas_Exp)


colnames(Table_Smoke_Gas_Exp) <- c("Smoke Gas Exposure", "Asthma Present (%)", "Mean FEV1", "Std Dev FEV1")

knitr::kable(Table_Smoke_Gas_Exp, align=c("l", "c", "c","c"), digits =2, caption = "Table 4. Forced expiratory volume in 1 second (mL) (FEV1) by Second Hand Smoke and Gas Stove Exposure")
```

## Data Visualization

# Generate color palette
```{r}
townfactor <- as.factor(hw2data$townname)

pal1 <- colorFactor(scale_fill_brewer("Pastel1"), hw2data$townname)

head(hw2data)

    
    numcolor = length(unique(hw2data$townname))
    mycolor1 <- c("red", "blue", "black", "brown", "purple", "green", "darkblue", "darkgreen", "orange", "maroon", "yellow", "cyan")
    pal1 <- mycolor1[1:numcolor] 
    
pal1 <- colorFactor(palette = c("red", "blue", "black", "brown", "purple", "green", "darkblue", "darkgreen", "orange", "maroon", "yellow", "cyan"), domain = hw2data$townname) 

```


#works except for legend
```{r}
#Plot 1
hw2data%>% 
  ggplot(mapping = aes(x = bmi_imp, y = fev_imp)) + 
  geom_point(mapping = aes(color = townname)) + 
  scale_fill_brewer(palette="Pastel1")+
  geom_smooth(method = lm, mapping = aes(linetype = townname)) +
  facet_wrap(~ townname, nrow = 3) +
  labs(title="Forced Expiratory Volume by Body Mass Index", x="Body Mass Index (kg/m2)", y= "Forced Expiratory Volume in 1 second (mL)") +
  
  addLegend('topright', pal = ~pal1, values=hw2data$townname, title='Town Name', opacity=1)
```
```{r}
#Plot 2 
Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.


hw2data %>%
  ggplot(mapping =aes(x=fev_imp, fill = obesity_level, color = obesity_level)) +   
  geom_histogram(bin = 100, binwidth = 50)+
  labs(title="Forced Expiratory Volume by Body Mass Index", x="Forced Expiratory Volume in 1 second (mL)", y= "Observations")+
scale_color_brewer(palette = "Greens")

```

```{r}
#Plot 3
hw2data %>%
  ggplot(mapping =aes(x=fev_imp, fill = smoke_gas_exposure, color = smoke_gas_exposure)) +   
  geom_histogram(bin = 100, binwidth = 50)+
  labs(title="Forced Expiratory Volume by Second Hand Smoke & Gas Stove Exposure", x="Forced Expiratory Volume in 1 second (mL)", y= "Observations")+
scale_color_brewer(palette = "Greens")
```
```{r}
#Plot 4

hw2data %>%
  ggplot(mapping =aes(x=smoke_gas_exposure, y=bmi_imp)) +   
  geom_bar()+
  labs(title="Body Mass Index by Second Hand Smoke and Gas Stove Exposure", x="Second Hand Smoke and Gas Stove Exposure", y= "Body Mass Index (kg/m2)")+
scale_color_brewer(palette = "Greens")
```

