---
title: "Homework 2"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(tidyverse)
library(dtplyr)
```
#Load in data from github
```{r}
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "individual.gz", method="libcurl", timeout = 60)

individual <- data.table::fread("individual.gz")

download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "regional.gz", method="libcurl", timeout = 60)

regional <- data.table::fread("regional.gz")
  
```
```{r}
dim(individual)
head(individual)
```
```{r}
head(regional)
#Trim down this dataset so it just has the info we are going to use (PM2.5 data and location for the leaflet)
regional_trim <- regional[, c("townname", "pm25_mass", "lon", "lat")]
regional_trim

```
#How many unique townnames are there? 
```{r}
unique(individual$townname)
unique(regional_trim$townname)
```

#Check variables for merging. There are 12 unique townnames. This matches with the 12 rows from the regional dataset. 
#Merge data
```{r}
hw2data <-merge(individual, regional_trim, by="townname" )
dim(hw2data)
```

#_This looks correct, there are 1200 rows and now the three additional columns added in the regional (trimmed) dataset
```{r}
head(hw2data)
tail(hw2data)
```
#Which variables have missing values?
```{r}
colSums(is.na(hw2data))
```


#For missing values, impute data using the average within the variables “male” and “hispanic.” For variables with 0/1 values will impute with the median.

```
```{r}
#Impute variables with missing values using mean or mode grouped by sex/hispanic

hw2data[, bmi_imp := fcoalesce(bmi, mean(bmi, na.rm = T)),
    by = .(male, hispanic)]


hw2data[, fev_imp := (fcoalesce(fev, mean(fev, na.rm = T))),
    by = .(male, hispanic)]

#Create another smoke and gasstove variable that is numeric so it can be imputed

hw2data [ , smoke_num := (as.numeric(smoke, na.rm=F))]
hw2data [ , gasstove_num := (as.numeric(gasstove, na.rm=F))]

#Now try imputation again

hw2data[, smoke_imp := (fcoalesce(smoke_num, median(smoke_num, na.rm = T))),
    by = .(male, hispanic)]

hw2data[, gasstove_imp := (fcoalesce(gasstove_num, median(gasstove_num, na.rm = T))),
    by = .(male, hispanic)]


```

# This worked. Now my dataset has the imputed variables in place. 


