---
title: "lab11"
author: "CB"
date: "`r Sys.Date()`"
output: html_document
link-citations: yes
editor_options: 
  chunk_output_type: console
always_allow_html: true
---


```{r setup,message=TRUE, warning=FALSE, echo=FALSE} 

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache=FALSE,
  include=TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```
```{r}
library(data.table)
library(dplyr)
library(tidyverse)
library(dtplyr)
library(knitr)
library(plotly)
library(widgetframe)
```

### Load both datasets 
```{r}

# load COVID state-level data from NYT
cv_states <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"))

# load state population data
state_pops <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/COVID19Tracking/associated-data/master/us_census_data/us_census_2018_population_estimates_states.csv"))
```

```{r, rename variables in population data}
state_pops$abb <- state_pops$state
state_pops$state <- state_pops$state_name
state_pops$state_name <- NULL
```
```{r}
# Merge data sets
cv_states <- merge(cv_states, state_pops, by="state")
```

### 2. Look at the data

- Inspect the dimensions, `head`, and `tail` of the data
- Inspect the structure of each variables. Are they in the correct format?

```{r}
dim(cv_states)
head(cv_states)
tail(cv_states)
str(cv_states)
```

### 3. Format the data
```{r}
# format the date
cv_states$date <- as.Date(cv_states$date, format="%Y-%m-%d")

```

```{r}
# format the state and state abbreviation (abb) variables
state_list <- unique(cv_states$state)
cv_states$state <- factor(cv_states$state, levels = state_list)
abb_list <- unique(cv_states$abb)
cv_states$abb <- factor(cv_states$abb, levels = abb_list)
```

```{r}
# order the data first by state, second by date
cv_states = cv_states[order(cv_states$state, cv_states$date),]
```

```{r}

# Confirm the variables are now correctly formatted
str(cv_states)
head(cv_states)
tail(cv_states)

# Inspect range of values for each variable. What is the date range? The range of cases and deaths?
head(cv_states)
summary(cv_states)
min(cv_states$date)
max(cv_states$date)
```
```{r}
# Add variables for new_cases and new_deaths


for (i in 1:length(state_list)) {
  cv_subset = subset(cv_states, state == state_list[i])
  cv_subset = cv_subset[order(cv_subset$date),]
  
  # add starting level for new cases and deaths
  cv_subset$new_cases = cv_subset$cases[1]
  cv_subset$new_deaths = cv_subset$deaths[1]
  

  for (j in 2:nrow(cv_subset)) {
    cv_subset$new_cases[j] = cv_subset$cases[j] - cv_subset$cases[j-1]
    cv_subset$new_deaths[j] = cv_subset$deaths[j] - cv_subset$deaths[j-1]
  }
  
  # include in main dataset
  cv_states$new_cases[cv_states$state==state_list[i]] = cv_subset$new_cases
  cv_states$new_deaths[cv_states$state==state_list[i]] = cv_subset$new_deaths
}
```
```{r}
# Focus on recent dates
cv_states <- cv_states %>% dplyr::filter(date >= "2022-06-01")
```
```{r}
p1<-ggplot(cv_states, aes(x = date, y = new_cases, color = state)) + geom_line() + geom_point(size = .5, alpha = 0.5)
ggplotly(p1)
p1<-NULL # to clear from workspace
```
```{r}
p2<-ggplot(cv_states, aes(x = date, y = new_deaths, color = state)) + geom_line() + geom_point(size = .5, alpha = 0.5)
ggplotly(p2)
p2<-NULL # to clear from workspace
```
```{r}
# set negative new case or death counts to 0
cv_states$new_cases[cv_states$new_cases<0] = 0
cv_states$new_deaths[cv_states$new_deaths<0] = 0
```

```{r}
# Recalculate `cases` and `deaths` as cumulative sum of updated `new_cases` and `new_deaths`

for (i in 1:length(state_list)) {
  cv_subset = subset(cv_states, state == state_list[i])
  
  # add starting level for new cases and deaths
  cv_subset$cases = cv_subset$cases[1]
  cv_subset$deaths = cv_subset$deaths[1]
  
  ### FINISH CODE HERE
  for (j in 2:nrow(cv_subset)) {
    cv_subset$cases[j] = cv_subset$new_cases[j] + cv_subset$cases[j-1]
    cv_subset$deaths[j] = cv_subset$new_deaths[j] + cv_subset$deaths[j-1]
  }
  
  # include in main dataset
  cv_states$cases[cv_states$state==state_list[i]] = cv_subset$cases
  cv_states$deaths[cv_states$state==state_list[i]] = cv_subset$deaths
}
```
```{r}
library(zoo)

# Smooth new counts
cv_states$new_cases = zoo::rollmean(cv_states$new_cases, k=7, fill=NA, align='right') %>% round(digits = 0)

cv_states$new_deaths = zoo::rollmean(cv_states$new_deaths, k=7, fill=NA, align='right') %>% round(digits = 0)
```
```{r}
# Inspect data again interactively
p2<-ggplot(cv_states, aes(x = date, y = new_deaths, color = state)) + geom_line() + geom_point(size = .5, alpha = 0.5)
ggplotly(p2)
p2=NULL
```
### 5. Add additional variables

```{r}

# add population normalized (by 100,000) counts for each variable
cv_states$per100k =  as.numeric(format(round(cv_states$cases/(cv_states$population/100000),1),nsmall=1))
cv_states$newper100k =  as.numeric(format(round(cv_states$new_cases/(cv_states$population/100000),1),nsmall=1))
cv_states$deathsper100k =  as.numeric(format(round(cv_states$deaths/(cv_states$population/100000),1),nsmall=1))
cv_states$newdeathsper100k =  as.numeric(format(round(cv_states$new_deaths/(cv_states$population/100000),1),nsmall=1))
```

```{r}
# add a naive_CFR variable = deaths / cases
cv_states = cv_states %>% mutate(naive_CFR = round((deaths*100/cases),2))
```

```{r}
# create a `cv_states_today` dataset
cv_states_today = subset(cv_states, date==max(cv_states$date))
```

## II. Scatterplots

### 6. Explore scatterplots using `plot_ly()`
```{r}
# pop_density vs. cases
cv_states_today %>% 
  plot_ly(x = ~pop_density, y = ~cases, 
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5))
```
```{r}
# filter out "District of Columbia"
cv_states_today_filter <- cv_states_today %>% filter(state!="District of Columbia")

# pop_density vs. cases after filtering
cv_states_today_filter %>% 
  plot_ly(x = ~pop_density, y = ~cases, 
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5))
```
```{r}
# pop_density vs. deathsper100k
cv_states_today_filter %>% 
  plot_ly(x = ~pop_density, y = ~deathsper100k,
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5))

```
```{r}

# Adding hoverinfo
cv_states_today_filter %>% 
  plot_ly(x = ~pop_density, y = ~deathsper100k,
          type = 'scatter', mode = 'markers', color = ~state,
          size = ~population, sizes = c(5, 70), marker = list(sizemode='diameter', opacity=0.5),
          hoverinfo = 'text',
          text = ~paste( paste(state, ":", sep=""), 
                paste(" Cases per 100k: ", per100k, sep="") , 
                paste(" Deaths per 100k: ", deathsper100k, sep=""), sep = "<br>")) %>%
  layout(title = "Population-normalized COVID-19 deaths (per 100k) vs. population density for US states",
        yaxis = list(title = "Deaths per 100k"), 
         xaxis = list(title = "Population Density"),
         hovermode = "compare")
```


### 7. Explore scatterplot trend interactively using `ggplotly()` and `geom_smooth()`

- For `pop_density` vs. `newdeathsper100k` create a chart with the same variables using `gglot_ly()`
- Explore the pattern between $x$ and $y$ using `geom_smooth()`
  - Explain what you see. Do you think `pop_density` is a correlate of `newdeathsper100k`?

```{r}
### FINISH CODE HERE
p <- ggplot(cv_states_today_filter, aes(x=pop_density, y=deathsper100k, size=population)) + geom_point() + geom_smooth()
ggplotly(p)
```
### _There is perhaps a weak correlation between at higher population densities (from 750 to 1250 there is an upward linear appearing trend). However I doubt this correlation is very strong. The difference in new deaths per 100K is only 100/100,000 or 0.1% different between a population density near 0 and the highest measured, 1250. 


### 8. Multiple line chart

- Create a line chart of the `naive_CFR` for all states over time using `plot_ly()`
  - Use the zoom and pan tools to inspect the `naive_CFR` for the states that had an increase in September. How have they changed over time?
- Create one more line chart, for Florida only, which shows `new_cases` and `new_deaths` together in one plot. Hint: use `add_layer()`
  - Use hoverinfo to "eyeball" the approximate peak of deaths and peak of cases. What is the time delay between the peak of cases and the peak of deaths?
  

```{r}

# Line chart for naive_CFR for all states over time using `plot_ly()`
plot_ly(cv_states, x = ~date, y = ~naive_CFR, color = ~state, type = "scatter", mode = "lines")

# I looked for states that had increases in Sept.. really don't see any.


# Line chart for Florida showing new_cases and new_deaths together
cv_states %>% filter(state=="Florida") %>% plot_ly(x = ~date, y = ~new_cases, type = "scatter", mode = "lines") %>% 
  add_lines(x = ~date, y = ~new_deaths, type = "scatter", mode = "lines") 
```


### 9. Heatmaps

Create a heatmap to visualize `new_cases` for each state on each date greater than June 1st, 2021
- Start by mapping selected features in the dataframe into a matrix using the **tidyr** package function `pivot_wider()`, naming the rows and columns, as done in the lecture notes
- Use `plot_ly()` to create a heatmap out of this matrix. Which states stand out?
- Repeat with `newper100k` variable. Now which states stand out? 
- Create a second heatmap in which the pattern of `new_cases` for each state over time becomes more clear by filtering to only look at dates every two weeks 

```{r }
# Map state, date, and new_cases to a matrix
library(tidyr)
cv_states_mat <- cv_states %>% select(state, date, new_cases) %>% dplyr::filter(date>as.Date("2022-06-15"))
cv_states_mat2 <- as.data.frame(pivot_wider(cv_states_mat, names_from = state, values_from = new_cases))
rownames(cv_states_mat2) <- cv_states_mat2$date
cv_states_mat2$date <- NULL
cv_states_mat2 <- as.matrix(cv_states_mat2)
```
```{r}

# Create a heatmap using plot_ly()
plot_ly(x=colnames(cv_states_mat2), y=rownames(cv_states_mat2),
             z=~cv_states_mat2,
             type="heatmap",
             showscale=T)

```
```{r}
# Repeat with newper100k
cv_states_mat <- cv_states %>% select(state, date, newper100k) %>% dplyr::filter(date>as.Date("2022-06-15"))
cv_states_mat2 <- as.data.frame(pivot_wider(cv_states_mat, names_from = state, values_from = newper100k))
rownames(cv_states_mat2) <- cv_states_mat2$date
cv_states_mat2$date <- NULL
cv_states_mat2 <- as.matrix(cv_states_mat2)


plot_ly(x=colnames(cv_states_mat2), y=rownames(cv_states_mat2),
             z=~cv_states_mat2,
             type="heatmap",
             showscale=T)
```
```{r}
# Create a second heatmap after filtering to only include dates every other week
filter_dates <- seq(as.Date("2022-06-15"), as.Date("2022-11-01"), by="2 weeks")

cv_states_mat <- cv_states %>% select(state, date, newper100k) %>% filter(date %in% filter_dates)
cv_states_mat2 <- as.data.frame(pivot_wider(cv_states_mat, names_from = state, values_from = newper100k))
rownames(cv_states_mat2) <- cv_states_mat2$date
cv_states_mat2$date <- NULL
cv_states_mat2 <- as.matrix(cv_states_mat2)

# Create a heatmap using plot_ly()
plot_ly(x=colnames(cv_states_mat2), y=rownames(cv_states_mat2),
             z=~cv_states_mat2,
             type="heatmap",
             showscale=T)

```

### 10. Map

- Create a map to visualize the `naive_CFR` by state on October 15, 2021
- Compare with a map visualizing the `naive_CFR` by state on most recent date
- Plot the two maps together using `subplot()`. Make sure the shading is for the same range of values (google is your friend for this)
- Describe the difference in the pattern of the CFR.

```{r}
### For specified date
pick.date = "2022-10-15"

# Extract the data for each state by its abbreviation
cv_per100 <- cv_states %>% 
  filter(date==pick.date) %>%
  select(state, abb, newper100k, cases, deaths) 
# select data
cv_per100$state_name <- cv_per100$state
cv_per100$state <- cv_per100$abb
cv_per100$abb <- NULL


# Create hover text
cv_per100$hover <- with(cv_per100, 
                        paste(state_name, '<br>', 
                    "Cases per 100k: ", newper100k, '<br>',
                    "Cases: ", cases, '<br>', 
                    "Deaths: ", deaths))

```
```{r}
# Set up mapping details
set_map_details <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)


# Make sure both maps are on the same color scale
shadeLimit <- 35


# Create the map
fig <- plot_geo(cv_per100, locationmode = 'USA-states') %>% 
  add_trace(
    z = ~newper100k, 
    text = ~hover, 
    locations = ~state,
    color = ~newper100k, 
    colors = 'Purples'
  )
fig <- fig %>% 
  colorbar(title = "Cases per 100k", limits = c(0,shadeLimit))

fig <- fig %>% 
  layout(
    geo = set_map_details
  )
fig_pick.date <- fig


```

### Map for today's date
```{r}
# Extract the data for each state by its abbreviation
cv_per100 <- cv_states_today %>%  
  select(state, abb, newper100k, cases, deaths) 
# select data
cv_per100$state_name <- cv_per100$state
cv_per100$state <- cv_per100$abb
cv_per100$abb <- NULL

# Create hover text
cv_per100$hover <- with(cv_per100, 
            paste(state_name, '<br>', 
            "Cases per 100k: ", newper100k, '<br>',
            "Cases: ", cases, '<br>', 
            "Deaths: ", deaths))

# Create the map
fig <- plot_geo(cv_per100, locationmode = 'USA-states') %>% 
  add_trace(
    z = ~newper100k, 
    text = ~hover, 
    locations = ~state,
    color = ~newper100k, 
    colors = 'Purples'
  )
fig <- fig %>% 
  colorbar(title = "Cases per 100k", limits = c(0,shadeLimit))
fig <- fig %>% 
  layout(
    geo = set_map_details
  )
fig_Today <- fig
fig_Today
```


```{r}
### Plot together 
finalfig <- subplot(fig_pick.date, fig_Today, nrows = 2) %>% 
  layout(showlegend = FALSE,
         title = paste('Cases per 100k by State', 
                       '<br>(Hover for value)'),
         hovermode = TRUE
         ) %>%
  colorbar(title = "Cases per 100k", limits = c(0,shadeLimit))
annotations = list( 
  list( 
    x = 0.5,
    y = 0.5,
    text = pick.date,  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE   
  ),  
  list( 
    x = 0.5,
    y = -0.05,
    text = Sys.Date(),  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))
finalfig <- finalfig %>%layout(annotations = annotations) 
finalfig
```
### _ Between Oct 15th, 2022 and Nov 14th, 2022 there has been a shift in the location of high cases per 100K. In October, Kentucky had the highest case rate. Now in November it is New Mexico. It appears that all the four corners states have more cases in Nov than they did in Oct. Some states however such as MOntana and Florida have a lower case rate in Nov than in Oct. In general New England region appears to have lower cases in Nov than they did in Oct._





















