---
title: "Lab 04"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(tidyverse)
```


## Step 1 Read in the data from last week's lab
```{r}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}
met <- data.table::fread("met_all.gz")

```
## Step 2 Prep data -- Remove Temps less than -17C and check data with head command

```{r}
met <- met[temp>-17]
met2 <- met[order(temp)]
head(met2)
```
# Replace missing values 9999 with “NA”

``` {r}
met[met$elev==9999.0] <- NA
summary(met$elev)
```
# Generate a date variable 

```{r create-ymd, cache=TRUE}

met <- met[ ,ymd :=as.Date(paste(year, month, day, sep = "-"))]

```
## Keep observations from first week of the month

```{r}


met[, table(week(ymd))]
met <- met[ week(ymd) == 31 ]
```


## Check that max values are true (and not 9999)
```{r check max}
met[, .(
  temp      = max(temp,na.rm=T),
  rh        = max(rh,na.rm=T), 
  wind.sp   = max(wind.sp,na.rm=T), 
  vis.dist  = max(vis.dist,na.rm=T), 
  dew.point = max(dew.point,na.rm=T), 
  lat       = max(lat,na.rm=T), 
  lon       = max(lon,na.rm=T), 
  elev      = max(elev,na.rm=T)
)]
```
## Find averages for each variable by station
```{r check avg}
met_avg <- met[, .(
  temp      = mean(temp,na.rm=T),
  rh        = mean(rh,na.rm=T), 
  wind.sp   = mean(wind.sp,na.rm=T), 
  vis.dist  = mean(vis.dist,na.rm=T), 
  dew.point = mean(dew.point,na.rm=T), 
  lat       = mean(lat,na.rm=T), 
  lon       = mean(lon,na.rm=T), 
  elev      = mean(elev,na.rm=T)
), by = "USAFID"]
```

## Create region variables 
```{r}
met_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW","SE")))
    ]
table(met_avg$region)

```
## Create categorical variable for elevation
```{r}
met_avg[, elev_cat := fifelse(elev > 252, "high", "low")]
```



## Make Violin plots - dewpoint
```{r}
met_avg[!is.na(region)] %>%
  
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = dew.point,  color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 1)
```
# _southeast has highest dewpoint_

# Violin plot wind speed
```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot() + 
  geom_violin(mapping = aes(x = 1, y = wind.sp,  color=region, fill = region)) + 
  facet_wrap(~ region, nrow = 2)

```
# _highest wind speeds in the Southwest with a bimodal distribution_

#Step 4 Examine association between dew point temperature and wind speed by region
```{r}
met_avg[!is.na(region) & !is.na(wind.sp)] %>% 
  ggplot(mapping = aes(x = wind.sp, y = dew.point)) + 
  geom_point(mapping = aes(color = region)) + 
  geom_smooth(method = lm, mapping = aes(linetype = region)) +
  facet_wrap(~ region, nrow = 2)
```
# _Only in the southwest is the dewpoint decreasing as wind speed increases_


##Step 5 Use geom_bar to create barplots of the weather stations by elevation category coloured by region
```{r}
met_avg[!is.na(elev_cat) & !is.na(region)]%>% #Remove missing values for these variables
ggplot(met_avg, mapping=aes(x=region, fill=elev_cat))+
  geom_bar(color="black", position="dodge") + scale_fill_brewer(palette="Greens")+
    labs(title="Weather Station Elevation by Region", x="Region", y="Number of Stations")
             
  
```
#The Southweast is the lowest elevation area by quite a bit. The Northwest has very minimal low elevation areas.


## Step 6 examined mean dew point and wind speed by region using SD error bars
```{r}
met_avg[!is.na(dew.point)] %>%
  ggplot(mapping = aes(x = region, y = dew.point)) + 
    stat_summary(fun.data = mean_sdl, geom = "pointrange") +
    stat_summary(fun.data = mean_sdl, geom = "errorbar") 
```

# There is a wider distribution of values for the West vs East sides of the US

## Step 7 Make map showing the spatial trend in relative humidity in the US
```{r}
library(leaflet)
```

# Generate color palette
```{r}
rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=met_avg$rh)
rh.pal
```


#Generate map
```{r}
top10rh <-met_avg[rank (-rh) <=10]
```
```{r}
rhmap <- leaflet(met_avg) %>% 

  addProviderTiles('CartoDB.Positron') %>% 

  addCircles(
    lat = ~lat, lng=~lon,
                                                
    label = ~paste0(rh), color = ~ rh.pal(rh),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%

  addLegend('bottomleft', pal=rh.pal, values=met_avg$rh,
          title='Relative Humidity', opacity=1)
rhmap
```


#It is more humid in the Eastern US vs Western. Perhaps the West coast humidity is areas like the Coastal Redwood forests?
