---
title: "Lab 05"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(data.table)
library(lubridate)
library(dplyr)
library(tidyverse)
library(dtplyr)
library(leaflet)
library(tidyr)
```

## Step 1 Read in the data from week 3 lab
```{r}
if (!file.exists("met_all.gz")) {
download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}
met <- data.table::fread("met_all.gz")

```
# Remove implausible and NA values from the met data again
```{r}
met <- met[temp>-17]
met[met$elev==9999.0] <- NA
```
## Step 2 Obtain station data and clean

# Download the data
```{r}
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]
```

# Dealing with NAs and 999999
```{r}
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]
```

# Selecting the three relevant columns, and keeping unique records
```{r}
stations <- unique(stations[, list(USAF, CTRY, STATE)])
```

# Dropping NAs
```{r}
stations <- stations[!is.na(USAF)]
```

# Removing duplicates
```{r}
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
head(stations, n = 4)

```

# Merge the two datasets using code from lecture
```{r}
met <-  
  merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) 

nrow(met)

```

```{r}
head(met)
```
## Find the representative station for the U.S.  
```{r}
#Find means 
station_averages <- 
     met[ , .(
       temp      = mean(temp, na.rm=T),
       wind.sp   = mean(wind.sp, na.rm=T),
       atm.press = mean(atm.press,na.rm=T)
     ), by = USAFID]


#Find medians
stmeds <- station_averages[ , .(
          temp50     = median(temp, na.rm=T),
          windsp50   = median(wind.sp,na.rm=T),
          atmpress50 = median(atm.press,na.rm=T))]

#View values
stmeds

```
```{r}
#Find the station w shortest distance using which.min function
station_averages[ , 
              temp_dist50 := abs(temp - stmeds$temp50)][order(temp_dist50)]

station_averages[ ,which.min(temp_dist50)]
```

#The station 720458 is the nearest one

## Find the station most representative by STATE
```{r}
#First get means
station_averages <- 
     met[ , .(
       temp      = mean(temp, na.rm=T),
       wind.sp   = mean(wind.sp, na.rm=T),
       atm.press = mean(atm.press,na.rm=T)
     ), by = .(USAFID,STATE)]
head(station_averages)
```

```{r}
#Now getting medians

statemeds<- station_averages[ , .(
           temp50    = median(temp, na.rm=T), 
           wind.sp50 = median(wind.sp, na.rm=T)
            ), by = STATE]
statemeds
```

#Merge data
```{r}
station_averages <- 
  merge(
  x = station_averages,
  y = statemeds,
  by.x = "STATE",
  by.y = "STATE",
  all.x = TRUE,
  all.y = FALSE
)
```

```{r}
#Find the minimum distance
station_averages[ , temp_dist_state50   := temp - temp50]
station_averages[ , windsp_dist_state50 := wind.sp - wind.sp50] 
station_averages
```
#Find the euclidian dist
```{r}
station_averages[ , eucdist := temp_dist_state50^2 +
                                windsp_dist_state50^2]
```
#Find the representative station
```{r}
repstation <- station_averages[ , .(
                    eucdist = min(eucdist, na.rm=T))
                  , by=STATE]
```

```{r}
repweather <- 
  merge(
  x = station_averages,
  y = repstation,
  by.x = c("eucdist","STATE"),
  by.y = c("eucdist","STATE"),
  all.x = FALSE,
  all.y = TRUE
)
repweather
```
#Chart shows stations with smallest euclidian distance


## Question 3  Identify the station closest to the mid-point of the state.

```{r}
#Find station locations
met_stations <- unique(met[, .(USAFID, STATE, lon, lat)])
nrow(met_stations)
```

```{r}
length(unique(met$USAFID))
```
#More station locations than station IDs

```{r}
#Choose stations from earliest recording in the month
met_stations <- unique(met[order(USAFID,day,hour)][, .(USAFID, STATE, lon, lat)])
met_stations[, n := 1:.N, by=USAFID]
met_stations <- met_stations[n == 1][, n:= NULL]
nrow(met_stations)
```

```{r}
met_stations[, lat_mid := quantile(lat, probs = .5, na.rm = TRUE), by=STATE]
met_stations[, lon_mid := quantile(lon, probs = .5, na.rm = TRUE), by=STATE]
# Looking at the euclidean distances
met_stations[, dist := sqrt((lat-lat_mid)^2+(lon-lon_mid)^2)]
met_stations[, mindist := which.min(dist), by=STATE]
```
```{r}
#Create an index for each state
met_stations[, n := 1:.N, by=STATE]
met_stations[order(STATE,dist)]
```

```{r}
#Save just shortest distance stations
met_location <- met_stations[n == mindist, .(USAFID, STATE, lon, lat)][order(STATE)]
met_location
```

```{r}
#Get station locations from question 2
repweather <-
  merge(
  x =   repweather[, .(USAFID, STATE)],
  y = met_stations[, .(USAFID, STATE, lat, lon)],
  by.x = c("USAFID","STATE"),
  by.y = c("USAFID","STATE"),
  all.x = TRUE,
  all.y = FALSE
) 

repweather[, station_type := "met"]
met_location <- met_location[, .(USAFID, STATE, lat, lon)][, station_type := "loc"]
all_stations <- rbind(repweather, met_location, use.names=TRUE)
```
```{r}
#Visualize data

ggplot(map_data("state"), aes(x = long, y = lat)) +
  geom_map(aes(map_id = region), map = map_data("state"), col = "lightgrey", fill = "gray") +
  geom_point(
    data    = all_stations,
    mapping = aes(x = lon, y = lat, col = station_type),
    inherit.aes = FALSE, cex = 2, alpha = 0.5
    )
```

```{r}
#Compute averages and classify
met[, state_temp := mean(temp, na.rm = TRUE), by = STATE]

met %>% 
  drop_na(c(state_temp))


met[, temp_cat := fifelse(
  state_temp < 20, "low-temp", 
  fifelse(state_temp < 25, "mid-temp", "high-temp"))
  ]
head(met)

```


```{r}
table(met$temp_cat, useNA = "always")
```
#All NAs removed
```{r}
tab <- met[, .(
  N_entries   = .N,
  N_stations  = length(unique(USAFID)),
  N_states    = length(unique(STATE)),
  avg_temp    = mean(temp, na.rm = TRUE),
  avg_wind.sp = mean(wind.sp, na.rm = TRUE),
  avg_atm.pre = mean(atm.press, na.rm = TRUE)
), by = temp_cat]

knitr::kable(tab)
```

