---
title: "Lab 06"
author: "CB"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, eval=FALSE, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
#Load required libraries
```{r}
library(tidytext)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(forcats)

```
#Load in data
```{r}
download.file("https://raw.githubusercontent.com//USCbiostats//data-science-data//master//00_mtsamples//mtsamples.csv", "mtsamples.csv", method="libcurl", timeout=60)
```
```{r}

mts<-read.csv("mtsamples.csv")
dim(mts)
str(mts)
mts<-as_tibble(mts)
mts
```
#Question 1: What specialties do we have and how many of each are there in the dataset?
We can use count() from dplyr"
```{r}
mts %>% count(medical_specialty)
```
#_The specialties are definitely overlapping. For instance there is surgery in addition to urology which is a surgical subspecialty. They are not evenly distributed, some categories have more than 1000 entries and others <10)_
```{r}
colnames(mts)
```

```{r}
specialties<-mts %>% count(medical_specialty)
specialties %>% arrange(desc(n))
```
```{r}
#Make a barplot of specialty counts
specialties %>% 
  top_n(10) %>% 
  ggplot (aes(x=n, y=fct_reorder(medical_specialty,n)))+
  geom_col()
```
# Tokenize in transcription column and count number of times top 20 words appear
```{r}
library(tidytext)
mts %>%
  unnest_tokens(word,transcription) %>% 
   count(word, sort=TRUE) %>% 
    top_n(20, n)
  
```
#_Lots of stop words, so we do not get much information from this._

## Question 3

#Remove stop words
```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>% 
  top_n(20, n) %>% 
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

#_Lots of numbers here_, remove these._

```{r}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  
  filter( !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

#_Now the numbers are gone, the words are much more useful to determine what text is about_

## Question 4

